### 2021-04-28 submission/Highlights.rmd
proof reading submission

---
title: "Highlights"
output:
    bookdown::pdf_document2: default
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left, pagewise]{lineno}
  - \usepackage{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Highlights
  - Standardized framework for risk-based assessment of treatment effect
    heterogeneity
  - Always run the framework after population-level effect estimation problems
  - Run federated analyses across different databases mapped to OMOP common data
    model