### 2021-08-26 submission/manuscript.rmd
more response work on reviewers 1 and 2

---
title: "A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases"
author: 
  - Alexandros Rekkas, MSc${^1}$ 
  - David van Klaveren, PhD$^{2,3}$, 
  - Patrick B. Ryan, PhD$^4$
  - Ewout W. Steyerberg, PhD$^{3,5}$
  - David M. Kent, PhD$^2$
  - Peter R. Rijnbeek, PhD${^1}$
output:
    bookdown::pdf_document2: default
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left, pagewise]{lineno}
  - \usepackage{caption}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jamia.csl
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(here)
library(kableExtra)
d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}
knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```
\thispagestyle{empty}
\vspace{8mm}

$^1$ Department of Medical Informatics, Erasmus University Medical Center,
Rotterdam, Netherlands

$^2$ Predictive Analytics and Comparative Effectiveness (PACE) Center, Institute
for Clinical Research and Health Policy Studies (ICRHPS), Tufts Medical Center,
Boston, MA, USA

$^3$ Department of Public Health, Erasmus University Medical Center, Rotterdam,
Netherlands

$^4$ Janssen Research and Development, 125 Trenton Harbourton Rd,
Titusville,NJ 08560, USA

$^5$ Department of Biomedical Data Sciences, Leiden University Medical Center,
Leiden, The Netherlands

\vspace{10mm}
**Corresponding author**
\singlespacing 
Alexandros Rekkas, MSc

Department of Medical Informatics

Erasmus University Medical Center

3000 CA Rotterdam, P.O. Box 2040

Email: a.rekkas@erasmusmc.nl
\onehalfspacing

\vspace{10mm}
**Funding**

This work has been performed in the European Health Data and
Evidence Network (EHDEN) project. This project has received funding from the
Innovative Medicines Initiative 2 Joint Undertaking (JU) under grant agreement
No 806968. The JU receives support from the European Union’s Horizon 2020
research and innovation programme and EFPIA.

\newpage
\newpage
# Abstract {-}
\singlespacing 
One of the aims of the Observational Health Data Sciences and Informatics
(OHDSI) initiative is population-level treatment effect estimation in large
observational databases. Since treatment effects are well-known to vary across
groups of patients with different baseline risk, we aimed to extend the OHDSI
library of open-source tools with a framework for risk-based assessment of
treatment effect heterogeneity. The proposed framework consists of five
steps: 1) definition of the problem, i.e. the population, the treatment, the
comparator and the outcome(s) of interest; 2) identification of relevant
databases; 3) development of a prediction model for the outcome(s) of
interest; 4) estimation of propensity scores within strata of predicted risk and
estimation of relative and absolute treatment effect within strata of predicted
risk; 5) evaluation and presentation of results. We demonstrate our framework by
evaluating heterogeneity of the effect of angiotensin-converting enzyme (ACE)
inhibitors versus beta blockers on a set of 9 outcomes of interest across three
observational databases. With increasing risk of acute myocardial infarction we
observed increasing absolute benefits, i.e. from -0.03% to 0.54% in the lowest
to highest risk groups. Cough-related absolute harms decreased from 4.1%
to 2.6%. The proposed framework may be useful for the evaluation of
heterogeneity of treatment effect on observational data that are mapped to the
OMOP Common Data Model. The proof of concept study demonstrates its feasibility
in large observational data. Further insights may arise by application to safety
and effectiveness questions across the global data network.

\vspace{10mm}

**Keywords**: observational data, heterogeneity of treatment effect, risk stratification, subgroup analysis
\newpage 
\doublespacing 
\linenumbers

# Introduction

Understanding how a treatment’s effect varies across patients---a concept
described as heterogeneity of treatment effects (HTE)---has been central to the
agenda for both personalized (or precision) medicine and comparative
effectiveness research. More formally, HTE has been defined as the non-random
variability in the direction or magnitude of a treatment effect, in which the
effect is measured using clinical outcomes [@KRAVITZ2004]. Usually, analyses
focus on the relative scale, where treatment effects are assessed one at a time
in patient subgroups defined from single covariates, an approach that suffers
from low power and multiplicity issues [@Yusuf1991; @FernandezyGarcia2010].
However, even with well-established constant relative effects, treatment benefit
(or harm) may vary substantially on the absolute scale.

\> > EDIT > >

***In recent years, a large number of methods has been developed for the
assessment of HTE, mainly in the RCT setting. Earlier work suggested separating
HTE analyses into exploratory, confirmatory, descriptive and predictive
[@Varadhan2013]. Exploratory analyses focus on hypothesis generation,
confirmatory analyses test subgroup effect hypotheses, descriptive analyses aim
at facilitating future synthesis of subgroup effects and predictive analyses
predict probabilities of benefit or harm in individual patients. Predictive HTE
approaches can be further subdivided into risk modeling, treatment effect
modeling and optimal treatment regime methods, based on the reference class used
for defining patient similarity when making individualized predictions or
recommendations [@Rekkas2020]. We focus on “risk modeling” approaches where
patients are divided into risk strata using either an existing or an internally
developed risk prediction model. Risk-stratum-specific estimates provide an
overview of the evolution of treatment effects with increasing risk both on the
relative and the absolute scale. Recently, systematic guidance on the
application of such methods has been developed [@Kent2010; @Kent2018].***

< < EDIT < <

\> > OLD > >

Predictive HTE analyses have been described as approaches that
provide predictions of potential outcomes in a particular patient with one
intervention versus an alternative, taking into account multiple relevant
patient characteristics [@Varadhan2013]. We focus on “risk modeling”
approaches where patients are divided into risk strata using either an existing
or an internally developed risk prediction model. Risk-stratum-specific
estimates provide an overview of the evolution of treatment effects with
increasing risk both on the relative and the absolute scale. Recently,
systematic guidance on the application of such methods has been developed
[@Kent2010; @Kent2018].

< < OLD < <

While these approaches were developed for application in randomized controlled
trials (RCTs), observational databases are also an appealing substrate.
Observational healthcare databases, such as administrative claims and electronic
health records, are already highly available for the analysis of
pharmacoepidemiologic research questions [@AdlerMilstein2017; @Dahabreh2014].
They are also often larger than many typical trials, providing excellent power
for HTE analysis, including heterogeneous populations. However, unlike RCTs,
treatment effects are subject to confounding, while the unique structure of
different databases calls for database-specific analysis plans that often are
not easily transportable.

\> > ADD > >

***Because of the latter issue, running analyses at scale demands a big investment
of time and effort, as researchers are forced to map their analysis plans to the
databases available to them.***

< < ADD < <

The Observational Health Data Sciences and Informatics (OHDSI) collaborative has
established an international network of data partners and researchers that aim
to bring out the value of health data through large-scale analytics by mapping
all available databases to the Observational Medical Outcomes Partnership (OMOP)
Common Data Model (CDM) [@hripcsak2015observational; @Overhage2012].

\> > EDIT > >

***We aimed to develop a framework for implementing risk-based assessment of
treatment heterogeneity in high-dimensional observational data, extending the
existing guidelines of the RCT setting. Our publicly available package provides
an out-of-the-box solution for implementing such analyses at scale within the
OHDSI network, taking advantage of the OMOP-CDM.***

< < EDIT < <

\> > OLD > >

We aimed to develop a framework for risk-based assessment of treatment effect heterogeneity
in high-dimensional observational data, which extends the existing guidelines of
the RCT setting.

< < OLD < <

We implemented the framework using existing OHDSI methods including the
patient-level prediction framework and the population-level effect estimation
framework based on new-user cohort design [@Reps2018; @Ryan2013]. As a
proof-of-concept we analyzed heterogeneity of the effects of first-line
hypertension treatment: we compared the effect of angiotensin converting enzyme
(ACE) inhibitors to beta blockers on 9 outcomes across three different US claims
databases.

# Materials and Methods

The proposed framework defines 5 distinct steps that enable a standardized
approach for risk-based assessment of treatment effect heterogeneity for
databases mapped to the OMOP-CDM. These are: 1) general definition of the
research aim; 2) identification of the database within which the analyses will
be performed; 3) a prediction step where internal or external prediction models
are used to assign patient-level risk predictions; 4) an estimation step where
absolute and relative treatment effects are estimated within risk strata; 5)
presentation and evaluation of the results. We developed an open-source
R-package for the implementation of the proposed framework and made it publicly
available. The source code can be found at
<https://github.com/OHDSI/RiskStratifiedEstimation>. An overview of the
procedure can be seen in Figure \@ref(fig:figure1).

```{r figure1, echo=FALSE, fig.cap="(A) Starting from a treatment (top), a comparator (bottom) and an outcome (middle) cohort we estimate the propensity scores on the entire target population. (B) We match patients on the propensity scores and estimate the prediction model. Since we match patients we develop the prediction model on smaller subset of the initial population and, therefore, the number of patients is smaller in B compared to A. (C) We apply the prediction model on the entire population (green: lower 25\\% of the risk distribution; yellow: patients with risk between 25\\% and 50\\% of the risk distribution; orange: patients with risk between 50\\% and 75\\% of the risk distribution; red: patients at risk higher than 75\\% of the risk distribution). (D) We separate in risk subgroups, here quarters. Within risk quarters propensity scores are estimated again and relative and absolute treatment effects are estimated.", out.width = '100%'}
# knitr::include_graphics(here("figures/Figure_1.tif"))
grid::grid.raster(tiff::readTIFF(here("figures/Figure_1.tif")))
```

## Step 1: General definition of the problem

The typical research aim is: "to compare the effect of treatment $T$ to a
comparator treatment $C$ in patients with disease $D$ with respect to outcomes
$O_1,\dots,O_n$".

\> > ADD > >

***Cohort definitions are crucial for this step of the framework. We define a
cohort as the set of patients who satisfy one or more inclusion criteria for a
duration of time. A cohort within the OHDSI setting is more than a set of
specific clinical codes, providing a definition of a logic for how to use that
code set. All cohort definitions consist of: an entry event, i.e. the time a
patient enters a cohort; a set of inclusion criteria applied to the initial
event cohort to further restrict the set of people, resulting in the
construction of the construction of the qualifying cohort; cohort exit criteria
that terminate the patient's presence in the cohort. Cohort definitions are
transportable, meaning they can be implemented in any database, provided that it
is mapped to the OMOP-CDM.***

< < ADD < <

  - A single treatment cohort ($T$) which includes patients with disease $D$
    receiving the target treatment of interest. For example, a set of
    hypertension patients within a database that receive angiotensin-converting
    enzyme inhibitors, followed from the time of initiation until the time of
    censoring.
  - A single comparator cohort ($C$) which includes patients with disease $D$
    receiving the comparator (control) treatment. For example, a set of patients
    in a database that receive beta blockers, followed from the time of
    initiation until the time of censoring.
  - One or more outcome cohorts ($O_1,\dots,O_n$) that contain patients
    developing the outcomes of interest. For example, the set of patients in a
    database that have at least one occurrence of acute myocardial infarction in
    their record.

## Step 2: Identification of the database

The aim of this step is the inclusion of databases that represent the patient
population of interest. It is required that the databases are mapped to the
OMOP-CDM. The inclusion of multiple databases potentially increases the
generalizability of results. Furthermore, the cohorts should preferably have
adequate sample size to ensure precise effect estimation, even within smaller
risk strata.

## Step 3: Prediction

We adopt the standardized framework for the generation of patient-level
prediction models using observational data that ensures adherence to existing
guidelines [@Collins2015; @Moons2015]. This prediction framework requires the
definition of two essential cohorts: a target cohort, i.e. a set of patients
that satisfy one or more inclusion criteria for a duration of time, and an
outcome cohort.

To generate the target cohort we pool the already defined treatment cohort and
comparator cohort. Further restrictions can be applied on the target cohort to
construct the final population on which the prediction model will be developed
(e.g. exclude patients with a prior outcome in their history, before being
included in the target cohort). To avoid deferentially fitting the prediction
model to patients across treatment arms, thus introducing spurious interactions
with treatment [@Burke2014; @vanKlaveren2019], we develop the patient-level
prediction model in the propensity score-matched (1:1) subset of the population.

< < MOVE: ps_matching > >

\> > ADD > >

***Other methods of fitting the propensity scores, such as random forest and
others (REF) can also be considered.***

< < ADD < <

Finally, we need to define the time horizon within which we aim to make
predictions and we also need to select the machine-learning algorithm we want to
use to generate patient-level predictions. Currently, the available options are
regularized logistic regression, random forest, gradient boosting machines,
decision tree, naive Bayes, K-nearest neighbors, neural network and deep
learning (convolutional neural networks, recurrent neural network and deep
nets).

## Step 4: Estimation

The aim of this step is the estimation of treatment effects (both on the
relative and the absolute scale) within risk strata---typically 4 risk
quarters---defined using the prediction model of step 3. Any appropriate method
for the evaluation of relative and absolute treatment effects can be considered,
as long as the this is done consistently in all risk strata. Common approaches
are odds ratios or hazard ratios for relative scale estimates and differences in
observed proportions or differences in Kaplan-Meier estimates for absolute scale
estimates, depending on the problem at hand. We estimate propensity scores
within risk strata which we then use to match patients from different treatment
cohorts or stratify them into groups with similar propensity scores or to weigh
each patient's contribution to the estimation process [@Austin2011].

## Step 5: Evaluation and presentation of results

In each risk stratum we need to evaluate if adequate covariate balance was
achieved accounting for measured confounding. If that is not the case,
interpretation of the results may be problematic. Common approaches include
evaluation of the overlap of propensity score distributions and calculation of
standardized covariate differences before and after propensity score adjustment.

A performance overview of the derived prediction models, including
discrimination and calibration both in the propensity score matched subset, the
entire population and separately for treated and comparator patients should also
be reported. This is quite relevant as the performance of the prediction models
is directly related to our ability to single out patient subgroups where
treatment may be highly beneficial or unsafe. Kent et al [@Kent2016]
demonstrated that the event rate and the discriminative ability of the
prediction model can predict very well the distribution of predicted risk. Lower
event rate and higher c-statistic (given good calibration) result in high risk
heterogeneity, thus making estimated average treatment effects uninformative. In
this case, risk stratified analysis of HTE can be more effective in singling out
patient subgroups that stand to benefit (or be harmed) most by treatment in
question.

In the presence of a positive treatment effect and a well-discriminating
prediction model we expect an increasing pattern of the differences in the
absolute scale, even if treatment effects remain constant on the relative scale
across risk strata. Results should be presented side-by-side on a
forest-plot-like format, so that the evolution of treatment effects across risk
strata is visible both on the relative and the absolute scale.

# Results

As a proof of concept, we focus on the comparison of angiotensin converting
enzyme (ACE) inhibitors to beta blockers are among the most common treatment
classes for hypertension, with well-established effectiveness. Beta blockers,
even though initially widely used for the treatment of hypertension, more recent
trials and meta-analyses have cast doubt on their relative effectiveness
[@wiysonge2017beta]. As a result, newer US guidelines do not consider beta
blockers for initial treatment for hypertension while in the EU guidelines
combination with other antihypertensive treatments is recommended
[@Whelton2018; @Williams2018]. However, another meta-analysis suggested that
the efficacy profile of beta blockers is similar to other major treatment
classes in younger hypertensive patients and, thus, countries like Canada still
include them as a first-line treatment candidate [@Khan2006; @Rabi2020].

## Step 1: General definition of the problem

We consider the following research aim: "compare the effect of ACE-inhibitors
($T$) to the effect of beta blockers ($C$) in patients with established
hypertension ($D$) with respect to 9 outcomes ($O_1,\dots,O_9$)". The cohorts
are:

  - Treatment cohort: Patients receiving any drug within the ACE-inhibitor class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Comparator cohort: Patients receiving any drug within the beta blocker class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Outcome cohorts: We consider 3 main and 6 safety outcome cohorts. These are
    patients in the database with a diagnosis of: acute myocardial infarction
    (MI); hospitalization with heart failure; ischemic or hemorrhagic stroke
    (efficacy outcomes); hypokalemia; hyperkalemia; hypotension; angioedema;
    cough; abnormal weight gain (safety outcomes).

All cohort definitions were identical to the ones used in the multinational
study carried out within OHDSI that provided overall treatment effect estimates
comparing all anti-hypertensive drug classes with each other [@Suchard2019].
More information can be found in the supplementary material.

## Step 2: Identification of the databases

We used the following databases:

  - IBM MarketScan Medicare Supplemental Beneficiaries (MDCR): Represents health
    services of retirees (aged 65 or older) in the United States with primary or
    Medicare supplemental coverage through privately insured fee-for-service,
    point-of-service or capitated health plans. These data include adjudicated
    health insurance claims (e.g. inpatient, outpatient and outpatient
    pharmacy). Additionally, it captures laboratory tests for a subset of the
    covered lives.
  - IBM MarketScan Medicaid (MDCD): Adjudicated US health insurance claims for
    Medicaid enrollees from multiple states. It includes hospital discharge
    diagnoses, outpatient diagnoses and procedures and outpatient pharmacy
    claims as well as ethnicity and Medicare eligibility.
  - IBM MarketScan Commercial Claims and Encounters (CCAE): Data from
    individuals enrolled in US employer-sponsored insurance health plans. The
    data includes adjudicated health insurance claims (e.g. inpatient,
    outpatient, and outpatient pharmacy) as well as enrollment data from large
    employers and health plans who provide private healthcare coverage to
    employees, their spouses and dependents. Additionally, it captures
    laboratory tests for a subset of the covered lives.

## Step 3: Prediction

We developed separate prediction models for all efficacy outcomes (acute MI,
hospitalization with heart failure and hemorrhagic or ischemic stroke) in each
database.

\> > MOVE: ps_matching > >

***More specifically, we first estimate propensity scores using LASSO
logistic regression and a large set of baseline covariates including
demographics, drug exposures, diagnoses, measurements and medical devices. We
match patients 1-1 using a caliper, i.e. the maximum distance that is acceptable
for any match. The default value we use is 0.2 on the standardized logit scale
for the propensity scores.***

< < MOVE < <

The prediction models were estimated on the propensity score matched (1:1) subset of
the population, using caliper of 0.2 and after excluding patients having the
outcomes any time prior to treatment initiation. We chose a time horizon of 2
years after inclusion into the target cohort. For this demonstration, we
developed the prediction models using LASSO logistic regression with 3-fold
cross validation for hyper-parameter selection. We chose a time horizon of 2
years after inclusion into the target cohort. For this demonstration, we
developed the prediction models using LASSO logistic regression with 3-fold
cross validation for hyper-parameter selection.

## Step 4: Estimation

We used patient-level predictions to stratify the patient population into 4 risk
quarters. Within risk strata, relative effects were estimated using Cox
regression and absolute effects were estimated from the Kaplan-Meier estimate
differences at 2 years after treatment initiation. To adjust for observed
confounding within risk strata, we estimated propensity scores using the same
approach as in the development of prediction models. We used the estimated
propensity scores to stratify patients into 5 strata, within each risk quarter.

## Step 5: Result presentation and evaluation

We present the results of the analysis in the CCAE database with
stratification based on risk predictions of acute MI. Results of analyses in the
other databases and with other risk stratifications are included in the
supplementary material.

For each outcome and in each risk stratum there were adequate numbers of
patients (Table \@ref(tab:table)). The discriminative ability of the prediction
models was moderate in the matched development subset (c-index 0.76 for acute MI
; 0.79 for hospitalization with heart failure; 0.74 for stroke; ), in the
general population (c-index 0.74 for acute MI; 0.77 for hospitalization with
heart failure; 0.73 for stroke), in the treatment cohort (c-index for acute MI
it was 0.71, for hospitalization with heart failure was 0.76 and for stroke it
was 0.72) and in the comparator cohort (c-index for acute MI it was 0.79 for
hospitalization with heart failure was 0.79 and for stroke it was 0.75).
\captionsetup{width=\textwidth}
```{r table, echo=FALSE, results='asis', booktabs=TRUE}
table <- tibble(
  outcome = c(
    "Acute myocardial infarction",
    rep("", 3),
    "Heart failure (hosp)",
    rep("", 3),
    "Stroke (ischemic or hemorrhagic)",
    rep("", 3)
  ),
  riskQuarter = rep(1:4, 3),
  patientsAce = c(
    161099, 204882, 214413, 204167,
    146259, 188006, 218052, 230226,
    146069, 187524, 217070, 226128
  ),
  personYearsAce = c(
    276171, 372197, 393583, 351727,
    249809, 341014, 399394, 400330,
    294484, 340234, 397830, 393861
  ),
  eventsAce = c(
    203, 534, 117, 2095,
    228, 457, 826, 2012,
    299, 554, 947, 1718
  ),
  patientsBeta = c(
    133977, 90193, 80662, 90908,
    126387, 84280, 83421, 98380,
    126264, 84000, 83038, 97628
  ),
  personYearsBeta = c(
    220633, 169231, 150035, 154419,
    206706, 158425, 155222, 169139,
    206453, 157913, 154587, 167810
  ),
  eventsBeta = c(
    135, 321, 535, 1520,
    378, 340, 570, 1773,
    320, 351, 521, 1077
  )
)

knitr::kable(
  table,
  format = "latex",
  longtable = TRUE,
  booktabs = TRUE,
  col.names = c(
    "Outcome", "Risk quarter",
    "Persons", "Person years", "Events",
    "Persons", "Person years", "Events"
  ),
   format.args = list(
    big.mar = ","
  ),
  caption = "Number of patients, person years and events within quarters of predicted risk for hospitalization with heart failure for the 3 main outcomes of the study (acute myocardial infarction, hospitalization with heart failure and ischemic or hemorrhagic stroke)."
) %>%
  kableExtra::add_header_above(
    c(
      "", "", 
      "ACE inhibitors" = 3,
      "Beta blockers"  = 3
    )
  )
```

Relative treatment effects of ACE-inhibitors vs beta blockers increased in favor
of ACE-inhibitors (hazard ratios decreased) with increasing acute MI risk,
resulting in more pronounced absolute risk differences (ARD) with increasing
acute MI risk (Figure \@ref(fig:figure2)).

```{r figure2, echo=FALSE, fig.cap="Overview of heterogeneity of ACE-inhibitors treatment within strata of predicted risk of acute MI. The top panel contains the observed acute rates of ACE-inhibitors and beta blockers within each quarter of predicted acute MI risk. These are derived using the KaplanMeier estimates at 730 days after inclusion. The middle panel, contains the hazard ratios of comparing ACE-inhibitors to beta blockers with regard to acute MI. These are estimated using Cox proportional hazards regression within quarters of predicted acute MI risk. The bottom panel contains absolute risk reduction for ACE-inhibitors compared to beta blockers. These are derived as the difference in Kaplan-Meier estimates at 730 after inclusion. Hazard ratios in the middle panel show a decreasing trend with increasing acute MI risk. Given the rather good discrimination of the prediction model (AUC=0.74), this results in an increasing trend for absolute benefit in favor of ACE-inhibitors with increasing risk.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/Figure_2.tif")))
```

In general, with increasing acute MI risk we observed an increasing pattern in
terms of absolute benefit for the main outcomes, while the absolute harms did
not increase for the safety outcomes. More specifically, patients in the low
risk quarter did not receive absolute benefit (-0.03%) while absolute risk
was 0.54% lower (95% confidence interval 0.36%—0.71%) for patients in the
highest risk quarter. In contrast, the absolute and relative effects of
ACE-inhibitors on safety outcomes (e.g. cough and angioedema) are slightly
decreasing with increasing acute MI risk (Figure \@ref(fig:figure3) and
\@ref(fig:figure4)). Similar results were observed in the other two databases
(see supplementary material).

```{r figure3, echo=FALSE, fig.cap="Hazard ratios (relative treatment effects) for the main and safety outcomes, estimated by fitting stratified Cox regression models within quarters of predicted risk of acute myocardial infarction (MI). The four risk quarters (Q1-Q4) are defined using the internally developed model for acute MI", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/Figure_3.tif")))
```

```{r figure4, echo=FALSE, fig.cap="Absolute risk reduction for the main and safety outcomes, estimated by fitting stratified Cox regression models within quarters of predicted risk of acute myocardial infarction (MI). The four risk quarters (Q1-Q4) are defined using the internally developed model for acute MI", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/Figure_4.tif")))
```

These results suggest that treatment with ACE-inhibitors, compared to treatment
with beta blockers, may be focused on the higher risk patients, in whom the
benefits outweigh the harms, while beta blockers may be a viable option in lower
risk patients, in whom the benefit-harm tradeoff is more favorable. This is in
accordance with earlier findings that beta blockers should also be considered as
first-line treatment for younger hypertensive patients [@Cruickshank2007;
@Khan2006]. More thorough evaluation of these results is required in future
research, however.

The results of the analyses performed can be accessed and assessed through a
publicly available web application (https://data.ohdsi.org/AceBeta9Outcomes).

# Discussion

We developed a framework for the assessment of heterogeneity of treatment effect
in large observational databases using a risk modeling approach. The framework
is implemented in an open source R-package in the OHDSI methods library
(<https://github.com/OHDSI/RiskStratifiedEstimation>). As a proof-of-concept, we
used our framework to evaluate heterogeneity of the effect of treatment with
ACE-inhibitors compared to beta blockers on 3 efficacy and 6 safety outcomes.

In recent years several methods for the evaluation of treatment effect
heterogeneity have been developed in the setting of RCTs [@Rekkas2020].
However, low power and restricted prior knowledge on the mechanisms of variation
in treatment effect are often inherent in RCTs, which are usually adequately
powered only for the analysis of the primary outcome. Observational databases
contain a large amount of information on treatment assignment and outcomes of
interest, while also capturing key patient characteristics. Our framework
provides a standardized approach that can be used to leverage available
information from these data sources, allowing for large-scale risk-based
assessment of treatment effect heterogeneity. Multiple outcomes can be evaluated
in patient subgroups of similar baseline outcome risk. Multiple outcome risk
stratification schemes can also be considered. However, this should be done with
caution, as it may hinder the interpretability of the results, in a similar
manner as typical subgroup analyses.

Recently, guidelines on the application of risk modeling approaches for the
assessment of heterogeneity of treatment effect in RCT settings have been
proposed [@Kent2019; @PathEnE]. Our framework aims to translate these
guidelines to the observational setting while also providing a toolkit for its
implementation within OHDSI. It encourages open science as it requires accurate
definition of the research questions translated into clear and reproducible
cohort definitions that can easily be shared among researchers. Researchers with
access to different databases mapped to OMOP-CDM can also very easily extend
their overall analyses with risk-based assessment of treatment effect
heterogeneity. This enables collaboration among multiple sites with access to
different patient populations. We propose that the framework is implemented any
time treatment effect estimation in high-dimensional observational data is
undertaken.

Several considerations need to be made. First, estimates may be
biased due to the observational nature of the data. We attempt to account for
potential confounding by estimating propensity scores within strata of predicted
risk. These scores are estimated using regularized logistic regression on a
large set of pre-defined covariates. However, such approaches do not account for
unobserved confounding [@Liu2013]. Several sensitivity analyses have been
proposed in the literature for measuring the robustness of results in the
presence of unobserved confounding. Another approach is to calibrate estimates
and confidence intervals based on a large set of negative controls
[@Schuemie2014; @Schuemie2571]. Negative controls are treatment-outcome pairs
for which a null effect has been established. Estimating these effects within
available data provides an approximation of the null distribution that can be
used to empirically recalibrate effect estimates. Future work may extend our
framework with this type of analyses.

Our method provides a risk-stratified assessment of treatment effect
heterogeneity. However, even though stratification can provide a useful overview
for clinical interpretation, these results cannot be applied to individuals in a
straightforward manner, as we are still estimating subgroup effects
[@Kent2019]. Presentation of treatment effects as a continuous function of
risk would be more helpful, but is methodologically challenging. Future research
is necessary for the development of methods for continuous risk-based assessment
of HTE.

\> > EDIT > >

***Ideally, externally derived and adequately validated prediction model would
be preferred for analyzing treatment effect heterogeneity [@Kent2010]. In the
absence of such prediction models an internally-developed risk prediction model
can be considered. Earlier simulations of RCT studies have shown that internal
models developed on the combined treatment and control arms blinded to treatment
gave relatively unbiased estimates of treatment effect across the spectrum of
risk [@Burke2014].***

< < EDIT < <

\> > OLD > >

Externally derived and well transportable prediction models are preferred for
analyzing treatment effect heterogeneity [@Kent2010]. In the absence of such
prediction models, simulations of RCTs have shown that internal models can be
used to provide unbiased estimates of treatment effect across the spectrum of
baseline risk [@Burke2014].

< < OLD < <

However, in observational databases treatment arms
may significantly differ in sample size. Because the prediction model will
possibly better fit to the larger treatment arm, spurious treatment-covariate
interactions may be introduced in the prediction model, leading to sub-optimal
risk stratification. As a remedy, we first match the patients in the treatment
and the comparator cohorts on the basis of propensity scores. Additionally, we
propose to assess model performance in the separate treatment arms to evaluate
its aptness for risk stratification.

Recently, disease risk scores have been explored as an alternative to propensity
scores for balancing covariates [@Glynn2012; @Hansen2008]. In our method, the
objective of risk stratification is not balancing, but assessing the variation
of treatment effects on multiple outcomes across patients with different levels
of baseline risk. Although using the same risk model for balancing and
risk-based HTE analysis may sound attractive, we note that our method only uses
one risk model for stratification and one propensity score model for balancing,
while separate disease risk score models would be required to analyze treatment
effects for each of the multiple outcomes.

In conclusion, the proof-of-concept study demonstrates the feasibility of our
framework for risk-based assessment of treatment effect heterogeneity in large
observational data. The standardized framework is easily applicable and highly
informative whenever treatment effect estimation in high-dimensional
observational data is of interest. Our framework is a supplement to the
population-level effect estimation framework developed within OHDSI and, in the
presence of an adequately discriminating prediction model, can be used to make
the overall results more actionable for medical decision making.

\newpage
# References
\nolinenumbers
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent