### 2021-09-09 submission/manuscript.rmd
modify results section to demonstrate power of OMOP

---
title: "A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases"
author: 
  - Alexandros Rekkas, MSc${^1}$ 
  - David van Klaveren, PhD$^{2,3}$, 
  - Patrick B. Ryan, PhD$^4$
  - Ewout W. Steyerberg, PhD$^{3,5}$
  - David M. Kent, PhD$^2$
  - Peter R. Rijnbeek, PhD${^1}$
output:
    bookdown::pdf_document2: default
    bookdown::word_document2:
        reference_docx: reference.docx
geometry: margin=1.0in
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left, pagewise]{lineno}
  - \usepackage{caption}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jamia.csl
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(here)
library(kableExtra)
d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}
knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```
\thispagestyle{empty}
\vspace{8mm}

$^1$ Department of Medical Informatics, Erasmus University Medical Center,
Rotterdam, Netherlands

$^2$ Predictive Analytics and Comparative Effectiveness (PACE) Center, Institute
for Clinical Research and Health Policy Studies (ICRHPS), Tufts Medical Center,
Boston, MA, USA

$^3$ Department of Public Health, Erasmus University Medical Center, Rotterdam,
Netherlands

$^4$ Janssen Research and Development, 125 Trenton Harbourton Rd,
Titusville,NJ 08560, USA

$^5$ Department of Biomedical Data Sciences, Leiden University Medical Center,
Leiden, The Netherlands

\vspace{10mm}
**Corresponding author**
\singlespacing 
Alexandros Rekkas, MSc

Department of Medical Informatics

Erasmus University Medical Center

3000 CA Rotterdam, P.O. Box 2040

Email: a.rekkas@erasmusmc.nl
\onehalfspacing

\vspace{10mm}
**Funding**

This work has been performed in the European Health Data and
Evidence Network (EHDEN) project. This project has received funding from the
Innovative Medicines Initiative 2 Joint Undertaking (JU) under grant agreement
No 806968. The JU receives support from the European Union’s Horizon 2020
research and innovation programme and EFPIA.

\newpage
\newpage
# Abstract {-}
\singlespacing 
One of the aims of the Observational Health Data Sciences and Informatics
(OHDSI) initiative is population-level treatment effect estimation in large
observational databases. Since treatment effects are well-known to vary across
groups of patients with different baseline risk, we aimed to extend the OHDSI
library of open-source tools with a framework for risk-based assessment of
treatment effect heterogeneity. The proposed framework consists of five
steps: 1) definition of the problem, i.e. the population, the treatment, the
comparator and the outcome(s) of interest; 2) identification of relevant
databases; 3) development of a prediction model for the outcome(s) of
interest; 4) estimation of propensity scores within strata of predicted risk and
estimation of relative and absolute treatment effect within strata of predicted
risk; 5) evaluation and presentation of results. We demonstrate our framework by
evaluating heterogeneity of the effect of angiotensin-converting enzyme (ACE)
inhibitors versus beta blockers on a set of 9 outcomes of interest across three
observational databases. With increasing risk of acute myocardial infarction we
observed increasing absolute benefits, i.e. from -0.03% to 0.54% in the lowest
to highest risk groups. Cough-related absolute harms decreased from 4.1%
to 2.6%. The proposed framework may be useful for the evaluation of
heterogeneity of treatment effect on observational data that are mapped to the
OMOP Common Data Model. The proof of concept study demonstrates its feasibility
in large observational data. Further insights may arise by application to safety
and effectiveness questions across the global data network.

\vspace{10mm}

**Keywords**: observational data, heterogeneity of treatment effect, risk stratification, subgroup analysis
\newpage 
\doublespacing 
\linenumbers

# Introduction

Understanding how a treatment’s effect varies across patients---a concept
described as heterogeneity of treatment effects (HTE)---has been central to the
agenda for both personalized (or precision) medicine and comparative
effectiveness research. More formally, HTE has been defined as the non-random
variability in the direction or magnitude of a treatment effect, in which the
effect is measured using clinical outcomes [@KRAVITZ2004]. Usually, analyses
focus on the relative scale, where treatment effects are assessed one at a time
in patient subgroups defined from single covariates, an approach that suffers
from low power and multiplicity issues [@Yusuf1991; @FernandezyGarcia2010].
However, even with well-established constant relative effects, treatment benefit
(or harm) may vary substantially on the absolute scale.

\> > EDIT > >

***In recent years, a large number of methods has been developed for the
assessment of HTE, mainly in the RCT setting. Earlier work suggested separating
HTE analyses into exploratory, confirmatory, descriptive and predictive
[@Varadhan2013]. Exploratory analyses focus on hypothesis generation,
confirmatory analyses test subgroup effect hypotheses, descriptive analyses aim
at facilitating future synthesis of subgroup effects and predictive analyses
predict probabilities of benefit or harm in individual patients. Predictive HTE
approaches can be further subdivided into risk modeling, treatment effect
modeling and optimal treatment regime methods, based on the reference class used
for defining patient similarity when making individualized predictions or
recommendations [@Rekkas2020]. We focus on “risk modeling” approaches where
patients are divided into risk strata using either an existing or an internally
developed risk prediction model. Risk-stratum-specific estimates provide an
overview of the evolution of treatment effects with increasing risk both on the
relative and the absolute scale. Recently, systematic guidance on the
application of such methods has been developed [@Kent2010; @Kent2018].***

< < EDIT < <

\> > OLD > >

Predictive HTE analyses have been described as approaches that
provide predictions of potential outcomes in a particular patient with one
intervention versus an alternative, taking into account multiple relevant
patient characteristics [@Varadhan2013]. We focus on “risk modeling”
approaches where patients are divided into risk strata using either an existing
or an internally developed risk prediction model. Risk-stratum-specific
estimates provide an overview of the evolution of treatment effects with
increasing risk both on the relative and the absolute scale. Recently,
systematic guidance on the application of such methods has been developed
[@Kent2010; @Kent2018].

< < OLD < <

While these approaches were developed for application in randomized controlled
trials (RCTs), observational databases are also an appealing substrate.
Observational healthcare databases, such as administrative claims and electronic
health records, are already highly available for the analysis of
pharmacoepidemiologic research questions [@AdlerMilstein2017; @Dahabreh2014].
They are also often larger than many typical trials, providing excellent power
for HTE analysis, including heterogeneous populations. However, unlike RCTs,
treatment effects are subject to confounding, while the unique structure of
different databases calls for database-specific analysis plans that often are
not easily transportable.

\> > ADD > >

***Because of the latter issue, running analyses at scale demands a big investment
of time and effort, as researchers are forced to map their analysis plans to the
databases available to them.***

< < ADD < <

The Observational Health Data Sciences and Informatics (OHDSI) collaborative has
established an international network of data partners and researchers that aim
to bring out the value of health data through large-scale analytics by mapping
all available databases to the Observational Medical Outcomes Partnership (OMOP)
Common Data Model (CDM) [@hripcsak2015observational; @Overhage2012].

\> > EDIT > >

***We aimed to develop a framework for implementing risk-based assessment of
treatment heterogeneity in high-dimensional observational data, extending the
existing guidelines of the RCT setting. Our publicly available package provides
an out-of-the-box solution for implementing such analyses at scale within the
OHDSI network, taking advantage of the OMOP-CDM.***

< < EDIT < <

\> > OLD > >

We aimed to develop a framework for risk-based assessment of treatment effect heterogeneity
in high-dimensional observational data, which extends the existing guidelines of
the RCT setting.

< < OLD < <

We implemented the framework using existing OHDSI methods including the
patient-level prediction framework and the population-level effect estimation
framework based on new-user cohort design [@Reps2018; @Ryan2013]. As a
proof-of-concept we analyzed heterogeneity of the effects of first-line
hypertension treatment: we compared the effect of angiotensin converting enzyme
(ACE) inhibitors to beta blockers on 9 outcomes across three different US claims
databases.

# Materials and Methods

The proposed framework defines 5 distinct steps that enable a standardized
approach for risk-based assessment of treatment effect heterogeneity for
databases mapped to the OMOP-CDM. These are: 1) general definition of the
research aim; 2) identification of the database within which the analyses will
be performed; 3) a prediction step where internal or external prediction models
are used to assign patient-level risk predictions; 4) an estimation step where
absolute and relative treatment effects are estimated within risk strata; 5)
presentation and evaluation of the results. We developed an open-source
R-package for the implementation of the proposed framework and made it publicly
available. The source code can be found at
<https://github.com/OHDSI/RiskStratifiedEstimation>. An overview of the
procedure can be seen in Figure \@ref(fig:figure1).

```{r figure1, echo=FALSE, fig.cap="(A) Starting from a treatment (top), a comparator (bottom) and an outcome (middle) cohort we estimate the propensity scores on the entire target population. (B) We match patients on the propensity scores and estimate the prediction model. Since we match patients we develop the prediction model on smaller subset of the initial population and, therefore, the number of patients is smaller in B compared to A. (C) We apply the prediction model on the entire population (green: lower 25\\% of the risk distribution; yellow: patients with risk between 25\\% and 50\\% of the risk distribution; orange: patients with risk between 50\\% and 75\\% of the risk distribution; red: patients at risk higher than 75\\% of the risk distribution). (D) We separate in risk subgroups, here quarters. Within risk quarters propensity scores are estimated again and relative and absolute treatment effects are estimated.", out.width = '100%'}
# knitr::include_graphics(here("figures/Figure_1.tif"))
grid::grid.raster(tiff::readTIFF(here("figures/Figure_1.tif")))
```

## Step 1: General definition of the problem

The typical research aim is: "to compare the effect of treatment $T$ to a
comparator treatment $C$ in patients with disease $D$ with respect to outcomes
$O_1,\dots,O_n$".

\> > ADD > >

***Cohort definitions are crucial for this step of the framework. We define a
cohort as the set of patients who satisfy one or more inclusion criteria for a
duration of time. A cohort within the OHDSI setting is more than a set of
specific clinical codes, providing a definition of a logic for how to use that
code set. All cohort definitions consist of: an entry event, i.e. the time a
patient enters a cohort; a set of inclusion criteria applied to the initial
event cohort to further restrict the set of people, resulting in the
construction of the construction of the qualifying cohort; cohort exit criteria
that terminate the patient's presence in the cohort. Cohort definitions are
transportable, meaning that in theory they can be implemented in any database,
provided that it is mapped to the OMOP-CDM.***

< < ADD < <

  - A single treatment cohort ($T$) which includes patients with disease $D$
    receiving the target treatment of interest. For example, a set of
    hypertension patients within a database that receive angiotensin-converting
    enzyme inhibitors, followed from the time of initiation until the time of
    censoring.
  - A single comparator cohort ($C$) which includes patients with disease $D$
    receiving the comparator (control) treatment. For example, a set of patients
    in a database that receive beta blockers, followed from the time of
    initiation until the time of censoring.
  - One or more outcome cohorts ($O_1,\dots,O_n$) that contain patients
    developing the outcomes of interest. For example, the set of patients in a
    database that have at least one occurrence of acute myocardial infarction
    (MI) in their record.

## Step 2: Identification of the database

The aim of this step is the inclusion of databases that represent the patient
population of interest. It is required that the databases are mapped to the
OMOP-CDM. The inclusion of multiple databases potentially increases the
generalizability of results. Furthermore, the cohorts should preferably have
adequate sample size to ensure precise effect estimation, even within smaller
risk strata.

## Step 3: Prediction

We adopt the standardized framework for the generation of patient-level
prediction models using observational data that ensures adherence to existing
guidelines [@Collins2015; @Moons2015]. This prediction framework requires the
definition of two essential cohorts: a target cohort, i.e. a set of patients
that satisfy one or more inclusion criteria for a duration of time, and an
outcome cohort.

To generate the target cohort we pool the already defined treatment cohort and
comparator cohort. Further restrictions can be applied on the target cohort to
construct the final population on which the prediction model will be developed
(e.g. exclude patients with a prior outcome in their history, before being
included in the target cohort). To avoid deferentially fitting the prediction
model to patients across treatment arms, thus introducing spurious interactions
with treatment [@Burke2014; @vanKlaveren2019], we develop the patient-level
prediction model in the propensity score-matched (1:1) subset of the population.

< < MOVE: ps_matching > >

\> > ADD > >

***Other methods of fitting the propensity scores, such as random forest and
others (REF) can also be considered.***

< < ADD < <

Finally, we need to define the time horizon within which we aim to make
predictions and we also need to select the machine-learning algorithm we want to
use to generate patient-level predictions. Currently, the available options are
regularized logistic regression, random forest, gradient boosting machines,
decision tree, naive Bayes, K-nearest neighbors, neural network and deep
learning (convolutional neural networks, recurrent neural network and deep
nets).

After model development, a performance overview of the derived prediction models
including discrimination and calibration both in the propensity score matched
subset, the entire population and separately for treated and comparator patients
should also be reported. This is important to ensure that no overfitting of the
prediction model in one of the cohorts has occurred. In addition, the
performance of the prediction models is directly related to our ability to
single out patient subgroups where treatment may be highly beneficial or unsafe.
Kent et al [@Kent2016] demonstrated that the event rate and the discriminative
ability of the prediction model can predict very well the distribution of
predicted risk. Lower event rate and higher c-statistic (given good calibration)
result in high risk heterogeneity, thus making estimated average treatment
effects uninformative. In this case, risk stratified analysis of HTE can be more
effective in singling out patient subgroups that stand to benefit (or be harmed)
most by treatment in question.

## Step 4: Estimation

The aim of this step is the estimation of treatment effects (both on the
relative and the absolute scale) within risk strata---typically 4 risk
quarters---defined using the prediction model of step 3. Any appropriate method
for the evaluation of relative and absolute treatment effects can be considered,
as long as the this is done consistently in all risk strata. Common approaches
are odds ratios or hazard ratios for relative scale estimates and differences in
observed proportions or differences in Kaplan-Meier estimates for absolute scale
estimates, depending on the problem at hand. We estimate propensity scores
within risk strata which we then use to match patients from different treatment
cohorts or stratify them into groups with similar propensity scores or to weigh
each patient's contribution to the estimation process [@Austin2011].

Before focusing on the results of the estimation process we need to evaluate if
adequate covariate balance was achieved within each risk stratum accounting for
measured confounding. Common approaches include evaluation of the overlap of
propensity score distributions and calculation of standardized covariate
differences before and after propensity score adjustment.

## Step 5: Presentation of results

In the presence of a positive treatment effect and a well-discriminating
prediction model we expect an increasing pattern of the differences in the
absolute scale, even if treatment effects remain constant on the relative scale
across risk strata. Due to this scale-dependence of treatment effect
heterogeneity, results should be assessed both on the relative and the absolute
scale. We find that a side-by-side presentation on a forest-like format can give
a very good representation of our results.

# Results

As a proof of concept, we focus on the comparison of angiotensin converting
enzyme (ACE) inhibitors to beta blockers are among the most common treatment
classes for hypertension, with well-established effectiveness. Beta blockers,
even though initially widely used for the treatment of hypertension, more recent
trials and meta-analyses have cast doubt on their relative effectiveness
[@wiysonge2017beta]. As a result, newer US guidelines do not consider beta
blockers for initial treatment for hypertension while in the EU guidelines
combination with other antihypertensive treatments is recommended
[@Whelton2018; @Williams2018]. However, another meta-analysis suggested that
the efficacy profile of beta blockers is similar to other major treatment
classes in younger hypertensive patients and, thus, countries like Canada still
include them as a first-line treatment candidate [@Khan2006; @Rabi2020].

## Step 1: General definition of the problem

We consider the following research aim: "compare the effect of ACE-inhibitors
($T$) to the effect of beta blockers ($C$) in patients with established
hypertension ($D$) with respect to 9 outcomes ($O_1,\dots,O_9$)". The cohorts
are:

  - Treatment cohort: Patients receiving any drug within the ACE-inhibitor class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Comparator cohort: Patients receiving any drug within the beta blocker class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Outcome cohorts: We consider 3 main and 6 safety outcome cohorts. These are
    patients in the database with a diagnosis of: acute MI; hospitalization with
    heart failure; ischemic or hemorrhagic stroke (efficacy outcomes);
    hypokalemia; hyperkalemia; hypotension; angioedema; cough; abnormal weight
    gain (safety outcomes).

All cohort definitions were identical to the ones used in the multinational
study carried out within OHDSI that provided overall treatment effect estimates
comparing all anti-hypertensive drug classes with each other [@Suchard2019].
More information can be found in the supplementary material.

## Step 2: Identification of the databases

We used the following databases:

  - IBM MarketScan Medicare Supplemental Beneficiaries (MDCR): Represents health
    services of retirees (aged 65 or older) in the United States with primary or
    Medicare supplemental coverage through privately insured fee-for-service,
    point-of-service or capitated health plans. These data include adjudicated
    health insurance claims (e.g. inpatient, outpatient and outpatient
    pharmacy). Additionally, it captures laboratory tests for a subset of the
    covered lives.
  - IBM MarketScan Medicaid (MDCD): Adjudicated US health insurance claims for
    Medicaid enrollees from multiple states. It includes hospital discharge
    diagnoses, outpatient diagnoses and procedures and outpatient pharmacy
    claims as well as ethnicity and Medicare eligibility.
  - IBM MarketScan Commercial Claims and Encounters (CCAE): Data from
    individuals enrolled in US employer-sponsored insurance health plans. The
    data includes adjudicated health insurance claims (e.g. inpatient,
    outpatient, and outpatient pharmacy) as well as enrollment data from large
    employers and health plans who provide private healthcare coverage to
    employees, their spouses and dependents. Additionally, it captures
    laboratory tests for a subset of the covered lives.

Our analyses included a total of 784,561, 66,820 and 101,661 patients initiating
treatment with ACE inhibitors and 395,740, 45,999 and 69,798 patients initiating
treatment with beta blockers in CCAE, MDCD and MDCD respectively (Table
\@ref(tab:table)). Adequate numbers of patients were included in all strata of
predicted acute MI risk.

\captionsetup{width=\textwidth}
```{r table, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, results='asis', booktabs=TRUE}
table <- readr::read_csv(
  here::here("data/processed/table1.csv"),
  na = character()
) %>%
  select(-database)

knitr::kable(
  table,
  format = "latex",
  longtable = TRUE,
  booktabs = TRUE,
  col.names = c(
    "Outcome", "Risk quarter",
    "Persons", "Person years", "Events",
    "Persons", "Person years", "Events"
  ),
   format.args = list(
    big.mar = ","
  ),
  caption = "Number of patients, person years and events within quarters of predicted risk for acute MI for the 3 main outcomes of the study (acute MI, hospitalization with heart failure and ischemic or hemorrhagic stroke) across the 3 databases. The populations used for the evaluation of hosptialization with heart failure and stroke are subsets of the populations used for the evaluation of actute MI."
) %>%
  kableExtra::group_rows("CCAE", 1, 12) %>%
  kableExtra::group_rows("MDCD", 13, 24) %>%
  kableExtra::group_rows("MDCR", 25, 36) %>%
  kableExtra::add_header_above(
    c(
      "", "", 
      "ACE inhibitors" = 3,
      "Beta blockers"  = 3
    )
  )
```

## Step 3: Prediction

We internally developed separate prediction models for acute MI in all 3
databases.

\> > MOVE: ps_matching > >

***More specifically, we first estimate propensity scores using LASSO
logistic regression and a large set of baseline covariates including
demographics, drug exposures, diagnoses, measurements and medical devices. We
match patients 1-1 using a caliper, i.e. the maximum distance that is acceptable
for any match. The default value we use is 0.2 on the standardized logit scale
for the propensity scores.***

< < MOVE < <

The prediction models were estimated on the propensity score matched (1:1)
subset of the population, using caliper of 0.2 and after excluding patients
having the outcomes any time prior to treatment initiation. We chose a time
horizon of 2 years after inclusion into the target cohort. For this
demonstration, we developed the prediction models using LASSO logistic
regression with 3-fold cross validation for hyper-parameter selection. For this
demonstration, we developed the prediction models using LASSO logistic
regression with 3-fold cross validation for hyper-parameter selection.

The models developed in the 3 databases had moderate discriminative performance
(internally validated) with no major issues of overfitting to any cohort except
for the case of CCAE database in which the derived prediction model performed
better in the comparator cohort (Table \@ref(tab:tableAucs)). We also observed
lower performance of the prediction model developed in MDCR compared to the
other 2 databases. Results on the calibration of the prediction models can be
found in the supplement.

\captionsetup{width=\textwidth}
```{r tableAucs, cache=TRUE, echo=FALSE, results='asis', booktabs=TRUE}
table <- tibble(
  population = c(
    "Matched set",
    "Treatment cohort",
    "Comparator cohort",
    "Entire population"
  ),
  ccae = c(
    75.78, 70.65, 79.1, 74.06
  ),
  mdcd = c(
    78.12, 78.29, 80.26, 79.46
  ),
  mdcr = c(
    71.09, 69.03, 68.53, 69.07
  )
)

knitr::kable(
  table,
  format = "latex",
  longtable = TRUE,
  booktabs = TRUE,
  col.names = c(
    "Population", "CCAE", "MDCD", "MDCR"
  ),
   format.args = list(
    big.mar = ","
  ),
  caption = "Discriminiative ability (c-statistic) of the derived prediction models for acute myocardial infarction in the matched set (development set), the treatment cohort, the comparator cohort and the etnire population in CCAE, MDCD and MDCR"
)
```

## Step 4: Estimation

We used patient-level predictions to stratify the patient population into 4 risk
quarters. Within risk strata, relative effects were estimated using Cox
regression and absolute effects were estimated from the Kaplan-Meier estimate
differences at 2 years after treatment initiation. To adjust for observed
confounding within risk strata, we estimated propensity scores using the same
approach as in the development of prediction models. We used the estimated
propensity scores to stratify patients into 5 strata, within each risk quarter.

In general, there was sufficient overlap of propensity score distribution in all
risk strata (Figure \@ref(fig:psDensity)). If empirical equipoise was not
achieved, the validity of the comparative effectiveness estimates would be
questionable.

```{r psDensity, cache = TRUE, echo=FALSE, fig.cap="Preference score distributions for the evaluation of heterogeneity of the effect of ACE inhibitors compared to beta blockers on acute MI based on quarters of predicted acute MI risk. The preference score is a transformation of the propensity score that adjusts for prevalence differences between populations.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/PsDensity.tiff")))
```

Propensity score adjustment achieved balance for most of the considered
covariates, measured using standardized mean differences before and after
adjustment (Figure \@ref(fig:covariateBalance)). However, in lower risk strata
imbalances persisted for a substantial subset of the covariates, all related to
pregnancy findings. This was anticipated (use of ACE inhibitors is specifically
contraindicated during pregnancy) and was already pointed out in
[@Suchard2019].

```{r covariateBalance, cache = TRUE, echo=FALSE, fig.cap="Patient charactersistic balance for ACE inhibitors and beta blockers before and after stratification on the propensity scores. Each dot represents the standardized difference of means for a single covariate before (x-axis) and after (y-axis) stratification", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CovariateBalance.tiff")))
```

## Step 5: Presentation of results

For hospitalization with acute MI there was an increasing trend in favor ACE
inhibitors compared to beta blockers on the relative scale (hazard ratios
decreased) with increasing acute MI risk. More specifically, hazard ratios
decreased from 0.98 (0.77 to 1.26; 95% CI), 1.30 (0.51 to 3.22; 95% CI) and 1.03
(0.82 to 1.29; 95% CI) to 0.76 (0.71 to 0.82; 95% CI), 0.94 (0.82 to 1.07; 95%
CI) and 1.03 (0.93 to 1.15; 95% CI) in CCAE, MDCD and MDCR respectively (Figure
\@ref(fig:combinedRelative)). In terms of hospitalization with heart failure
relative treatment effect estimates favored ACE inhibitors across all risk
strata in all databases. We found no differences between the two treatments in
their effect on stroke on the relative scale. In terms of the safety outcomes we
found an increased risk of cough and angioedema on the relative scale across all
risk strata. In the case of cough, this effect decreased with increasing risk of
acute MI---from 1.37 (1.33 to 1.41; 95% CI), 1.35 (1.24 to 1.48; 95% CI)
and 1.37 (1.29 to 1.45; 95% CI) in the lowest acute MI risk quarter to 1.26
(1.22 to 1.29; 95% CI), 1.07 (1.00 to 1.14) and 1.10 (1.04 to 1.17; 95% CI) in
the highest acute MI risk quarter in CCAE, MDCD and MDCR respectively.

```{r combinedRelative, cache=TRUE, echo=FALSE, fig.cap="Overview of heterogeneity of ACE-inhibitors treatment on the relative scale (hazard ratios) within strata of predicted risk of acute MI. Values below 1 favor ACE inhibitors, while values above 1 favor beta blockers.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CombinedRelative.tiff")))
```

We observed an increasing trend of treatment effect on the absolute scale with
increasing acute MI risk in favor of ACE inhibitors in terms of acute MI in all
databases except for MDCR---from -0.03% (-0.08% to 0.01%; 95% CI), -0.05%
(-0.18% to 0.08%; 95% CI) and -0.02% (-0.24% to 0.19%; 95% CI) in the lowest
acute MI risk quarter to 0.54% (0.36% to 0.71%; 95% CI), 0.29% (-0.39%
to 0.97%; 95% CI) and -0.39% (-0.96% to 0.18%; 95% CI) in the highest acute MI
risk quarter in CCAE, MDCD and MDCR, respectively (Figure \@ref(fig:figure12)).
We found no difference on the absolute scale for stroke across risk strata.
Absolute risk differences did not favor ACE inhibitors compared to beta blockers
in terms of cough, even though this effect again diminished with increasing
acute MI risk---from -4.14% (-4.62% to -3.66%; 95% CI), -6.45% (-9.12% to
-3.78%; 95% CI) and -4.81% (-5.76% to -3.85%; 95% CI) in the lowest acute MI
risk quarter to -2.57% (-2.99% to -2.15%; 95% CI), -1.11% (-2.93% to 0.70%; 95%
CI) and -1.69% (-2.83% to -0.55%; 95% CI) in the highest acute MI risk quarter
in CCAE, MDCD and MDCR, respectively. In terms of angioedema absolute risk
differences were very small due to the rarity of the outcome.

```{r figure12, cache=TRUE, echo=FALSE, fig.cap="Overview of heterogeneity of ACE-inhibitors treatment on the absolute scale within strata of predicted risk of acute MI. Estimates of absolute treatment effect are derived as the difference in Kaplan-Meier estimates at 730 after inclusion. Values above 0 favor ACE inhibitors, while values below 0 favor beta blockers.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/CombinedAbsolute.tiff")))
```

These results suggest that treatment with ACE-inhibitors, compared to treatment
with beta blockers, may be focused on the higher risk patients, in whom the
benefits outweigh the harms, while beta blockers may be a viable option in lower
risk patients, in whom the benefit-harm tradeoff is more favorable. This is in
accordance with earlier findings that beta blockers should also be considered as
first-line treatment for younger hypertensive patients [@Cruickshank2007;
@Khan2006]. This analysis, however, was carried out as demonstration of the
framework and more rigorous analyses are required to make any suggestions for
clinical practice.

The results of the analyses performed can be accessed and assessed through a
publicly available web application (https://data.ohdsi.org/AceBeta9Outcomes).

# Discussion

We developed a framework for the assessment of heterogeneity of treatment effect
in large observational databases using a risk modeling approach. The framework
is implemented in an open source R-package in the OHDSI methods library
(<https://github.com/OHDSI/RiskStratifiedEstimation>). As a proof-of-concept, we
used our framework to evaluate heterogeneity of the effect of treatment with
ACE-inhibitors compared to beta blockers on 3 efficacy and 6 safety outcomes.

In recent years several methods for the evaluation of treatment effect
heterogeneity have been developed in the setting of RCTs [@Rekkas2020].
However, low power and restricted prior knowledge on the mechanisms of variation
in treatment effect are often inherent in RCTs, which are usually adequately
powered only for the analysis of the primary outcome. Observational databases
contain a large amount of information on treatment assignment and outcomes of
interest, while also capturing key patient characteristics. Our framework
provides a standardized approach that can be used to leverage available
information from these data sources, allowing for large-scale risk-based
assessment of treatment effect heterogeneity. Multiple outcomes can be evaluated
in patient subgroups of similar baseline outcome risk. Multiple outcome risk
stratification schemes can also be considered. However, this should be done with
caution, as it may hinder the interpretability of the results, in a similar
manner as typical subgroup analyses.

Recently, guidelines on the application of risk modeling approaches for the
assessment of heterogeneity of treatment effect in RCT settings have been
proposed [@Kent2019; @PathEnE]. Our framework aims to translate these
guidelines to the observational setting while also providing a toolkit for its
implementation within OHDSI. It encourages open science as it requires accurate
definition of the research questions translated into clear and reproducible
cohort definitions that can easily be shared among researchers. Researchers with
access to different databases mapped to OMOP-CDM can also very easily extend
their overall analyses with risk-based assessment of treatment effect
heterogeneity. This enables collaboration among multiple sites with access to
different patient populations. We propose that the framework is implemented any
time treatment effect estimation in high-dimensional observational data is
undertaken.

Several considerations need to be made. First, estimates may be
biased due to the observational nature of the data. We attempt to account for
potential confounding by estimating propensity scores within strata of predicted
risk. These scores are estimated using regularized logistic regression on a
large set of pre-defined covariates. However, such approaches do not account for
unobserved confounding [@Liu2013]. Several sensitivity analyses have been
proposed in the literature for measuring the robustness of results in the
presence of unobserved confounding. Another approach is to calibrate estimates
and confidence intervals based on a large set of negative controls
[@Schuemie2014; @Schuemie2571]. Negative controls are treatment-outcome pairs
for which a null effect has been established. Estimating these effects within
available data provides an approximation of the null distribution that can be
used to empirically recalibrate effect estimates. Future work may extend our
framework with this type of analyses.

Our method provides a risk-stratified assessment of treatment effect
heterogeneity. However, even though stratification can provide a useful overview
for clinical interpretation, these results cannot be applied to individuals in a
straightforward manner, as we are still estimating subgroup effects
[@Kent2019]. Presentation of treatment effects as a continuous function of
risk would be more helpful, but is methodologically challenging. Future research
is necessary for the development of methods for continuous risk-based assessment
of HTE.

\> > EDIT > >

***Ideally, externally derived and adequately validated prediction model would
be preferred for analyzing treatment effect heterogeneity [@Kent2010]. In the
absence of such prediction models an internally-developed risk prediction model
can be considered. Earlier simulations of RCT studies have shown that internal
models developed on the combined treatment and control arms blinded to treatment
gave relatively unbiased estimates of treatment effect across the spectrum of
risk [@Burke2014].***

< < EDIT < <

\> > OLD > >

Externally derived and well transportable prediction models are preferred for
analyzing treatment effect heterogeneity [@Kent2010]. In the absence of such
prediction models, simulations of RCTs have shown that internal models can be
used to provide unbiased estimates of treatment effect across the spectrum of
baseline risk [@Burke2014].

< < OLD < <

However, in observational databases treatment arms
may significantly differ in sample size. Because the prediction model will
possibly better fit to the larger treatment arm, spurious treatment-covariate
interactions may be introduced in the prediction model, leading to sub-optimal
risk stratification. As a remedy, we first match the patients in the treatment
and the comparator cohorts on the basis of propensity scores. Additionally, we
propose to assess model performance in the separate treatment arms to evaluate
its aptness for risk stratification.

Recently, disease risk scores have been explored as an alternative to propensity
scores for balancing covariates [@Glynn2012; @Hansen2008]. In our method, the
objective of risk stratification is not balancing, but assessing the variation
of treatment effects on multiple outcomes across patients with different levels
of baseline risk. Although using the same risk model for balancing and
risk-based HTE analysis may sound attractive, we note that our method only uses
one risk model for stratification and one propensity score model for balancing,
while separate disease risk score models would be required to analyze treatment
effects for each of the multiple outcomes.

In conclusion, the proof-of-concept study demonstrates the feasibility of our
framework for risk-based assessment of treatment effect heterogeneity in large
observational data. The standardized framework is easily applicable and highly
informative whenever treatment effect estimation in high-dimensional
observational data is of interest. Our framework is a supplement to the
population-level effect estimation framework developed within OHDSI and, in the
presence of an adequately discriminating prediction model, can be used to make
the overall results more actionable for medical decision making.

\newpage
# References
\nolinenumbers
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent